### 2020-07-02 text/content.tex
Edit after proofreading

Past research has analyzed the efficacy of social distancing interventions on reducing the spread of epidemics using the 1918 Spanish Flu in the U.S. \cite{Hatchett2007-gc,Markel2007-ad,Bootsma2007-ww} and seasonal viral infections in France \cite{Adda2016-mn}. Our knowledge of economic impacts, however, is limited \cite{Wren-Lewis2020-vc}. For this question, past data may be less relevant, as the importance of face-to-face communication has increased steadily in the last 100 years through urbanization \cite{Henderson2010-mv,Henderson2002-ji} and specialization increased in business services as well \cite{Herrendorf2014-jx,Duarte2019-kc}. 
Even if advances in information and communication technology made it increasingly possible to communicate with co-workers and customers without the need for physical face-to-face interactions, personal contacts are still inevitable in some industries. During the 2020 lockdown in the Netherlands only 31 percent of workers in the retail sector, 27 percent of the workers in the catering sector and 31 percent of the workers in the construction sector reported working from home \cite{von2020labour}.

@@

To measure business disruptions from social distancing, we turn to recent data on the task descriptions of each occupation \cite{National_Center_for_ONET_Development2020-wj}, the precise geographic location of non-farm businesses in the U.S. \cite{CBP}, and customer mobility patterns \cite{SafeGraph2020-gn}. We construct three groups of occupations. First, some occupations require face-to-face communication several times a week with other workers. Examples of these \emph{teamwork-intensive} occupations include maintenance, personal care related occupations and health care professionals. Other occupations require frequent face-to-face contact with customers. Retail salespersons, social workers and waiters and waitresses are examples of such \emph{customer-facing} occupations. The third group of workers may need to be in physical proximity of one another even if they do not communicate, for example, to operate machinery or access key resources. Examples of such occupations requiring \emph{physical presence} are drivers and machine operators, especially in mining and water transport, where crammed working environments are common. With our occupation level measures, we aim to capture the jobs that can be performed less efficiently from home. As O*NET measures the type of activities that are performed by the worker during normal times, we validate our indexes by using the American Time Use Survey (ATUS) \cite{ATUS}. The survey directly asks about the possibility of working from home.

To study how the patterns of interaction have changed in the U.S. during the Covid-19 pandemic, we use customer mobility data from SafeGraph \cite{SafeGraph2020-gn}. This dataset measures the number of visits to a business in a given month, as captured from several cell phone apps and made available to researchers in an anonymized form. ???????

@@

The optimal number of contacts without social distancing is given by Eq \eqref{nstar}. Firms with $n^*>N$ are limited by social distancing. Without moving communication online, their unit cost will increase to $c' = N \tau + N^{-\gamma}/\gamma$, which is greater than the optimal cost,
\begin{equation}\label{socdist}
\frac{c'}{c} 
	= \chi \frac{N}{n^*} + (1-\chi) \left(\frac{N}{n^*}\right)^{-\gamma}>1.
\end{equation}
The first term of the weighted average is less than one, representing a reduction in communication costs once the number of contacts declines. The second term is greater than one due to the fact that every worker has to complete a wider range of tasks than before, and they lose the benefit of specialization. Because $n^*$ is the cost-minimizing communication choice of the firm, the second term dominates, and production costs increase with social distancing.

@@

With our occupation-level measures, we aim to capture the type of activities that require frequent face-to-face contact. Our assumption is that these activities cannot be effectively done from home. To validate this assumption, we use data from the American Time Use Survey (ATUS) \cite{ATUS}, which directly asks the workers whether they \emph{can} work from home. Fig. \ref{customer_validation} plots our customer-intensity measure, for each occupation, against the share of workers who reported in ATUS that they can work from home (dark filled circles). Indeed, most customer-facing occupations have very few workers who can work from home. The pattern is different if we look at occupations that rely more on email, letters, and memos for customer communication (light hollow circles). The majority of these workers can work from home. Interestingly, for each degree of working from home (horizontal axis), there is sufficient variation in the importance of customer contact (vertical axis). The same patterns can be observed for teamwork-intensive occupations (Fig. \ref{teamwork_validation}). This makes use conclude that dropping occupations primarily relying on email, letters and memos is sufficient to control for the potential to work from home.

@@

Using the cost change in Eq \ref{socdist}, we can express the compensating labor subsidy of industry $i$ in region $r$ as
\begin{equation}\label{subsidy}
    \lambda_{ir} = 1 -
        \frac{1-\chi_i}{1-\chi_i N_{ir}/n_{ir}^*}
        \left(
        \frac{N_{ir}}{n_{ir}^*}
        \right)^{\gamma_i}>0.
\end{equation}
We calibrate $n_{ir}^*$ to match the number of customer visits to establishments of industry $i$ in ZIP-code $r$ in February 2020, assuming that these visits were optimal before the pandemic hit the U.S. The new number of visits, $N_{it}$ will be calibrated to the number of customer visits in May 2020. The compensating wage subsidy increases in the importance of communication $\chi_i$ and decreases in the change in cotacts $N_{ir}/n^*_{ir}$. The subscripts note that communication share is industry specific and the change in contacts is both industry and region specific.

@@

\begin{table}[!ht]
%\begin{adjustwidth}{-2.25in}{0in} % Comment out/remove adjustwidth environment if table fits in text column.
\caption{
{\bf Retail, accommodation and restaurants are the most communication intensive.}}
\begin{tabular}{|l+c|c|c+c|}
\hline
 & \multicolumn{3}{c}{\bf Communication} &  
\\ 
{\bf Industry} & {\bf Teamw.} & {\bf Custom.} & {\bf Overall} & {\bf Presence} 
\\ \thickhline
Retail trade &13&65&66&4\\
Accommodation \& food services &9&49&51&1\\
Arts, Entertainment, and Recreation &12&37&40&2\\
Other Services (except Public Admin.) &12&28&30&13\\
Admin. \& Support \& Waste Manag.  &	17&	24&	7&	27\\
\& Remed. Serv. &&&& \\
...&&&&\\
Information&	6&	13&	15&	9\\
Transportation and Warehousing&	8&	8&	14&	33\\
Prof., Scient., and Technical Serv.&	5&	10&	12&	1\\
Manufacturing&	7&	5&	9&	10\\
Agri., forestry, fishing \& hunting &	4&	1&	4&	23\\

@@

We see four avenues for further research. The first concerns the demand side of the economy, which we have mostly neglected by focusing on the production function. The employment response to a production disruption greatly depends on the elasticity of demand. We hypothesize that sectors like schooling and health care have inelastic demand and will continue to employ many workers despite significant disruptions. However, personal services, small grocery stores and restaurants may face more elastic demand and respond to large production cost increases by laying off a significant fraction of their work force.

The second direction concerns the interaction between sectors and regions. Whenever productivity in any business drops, this shock can propagate to its buyers and suppliers. The aggregate consequences of the epidemic will hence be modulated by input-output linkages between sectors, regions and countries \cite{Caliendo2014-mr,Caselli2020-nf,Baldwin2020-wb,Barrot2020-jf}.

The third and forth directions concern the long-run response of businesses as they try to become more resilient to such shocks in the future. Whether the share of telecommunication remains large in the long run depends crucially on how easily it substitutes for face-to-face interaction. In our model, the two are perfect substitutes with the only distinction that the efficiency of face-to-face meetings improves with population density, whereas telecommunication does not depend on agglomeration \cite{Rossi-Hansberg2007-tm,Ioannides2008-bs,Tian2019-wq}. Previous work has found face-to-face communication to be more effective in high-intensity communication which is particularly helpful to overcome incentive problems in joint production \cite{Gaspar1998-gy,Storper2004-mg}. Data on internet flows suggests that telecommunication is not a good substitute for face-to-face meetings \cite{Cuberes2013-js}. None of these papers discuss disruptions from social distancing measures. Further study of the modes of communication in the O*NET occupation survey can shed light on whether telecommunication can act as a low-cost substitute for face-to-face meetings.

Fourth, businesses may change their location in response to perceived threats and disruptions. As we discussed, epidemics have a disproportionate effect on cities. So it is conceivable that in a post-pandemic spatial equilibrium (not modeled here, but see \cite{Tian2019-wq}), the agglomeration premium falls and firms find it less attractive to locate in cities. A poignant point of comparison is the increased threat of terrorism in major cities following devastating attacks on New York, Washington, London, Paris, Madrid, Moscow and Mumbai. The general conclusion about terror threat is that cities have remained resilient and a robust attractor of businesses \cite{Glaeser2002-mw,Harrigan2002-ik}. We speculate that epidemics and social distancing can be more detrimental to cities than terror threats, because they tear apart the very fabric of urban life. However, we have limited data to make further predictions.

\section*{Data Availability Statement}
The mobility data used in this paper \cite{SafeGraph2020-gn} is confidential, but may be obtained free of charge for Covid-19-related research from the COVID-19 Consortium. Researchers interested in access to the data can apply at https://www.safegraph.com/covid-19-data-consortium. After signing a Data Agreement, access is granted within a few days. The authors will assist with any reasonable replication attempts for two years following publication.

@@

\section*{Acknowledgments}
We thank Gábor Békés, Péter Harasztosi, Péter Karádi, Balázs Lengyel, Dávid Krisztián Nagy, Andrea Weber, Charles Wyplosz, and an anonymous referee for comments.

>>>>

Past research has analyzed the efficacy of social distancing interventions on reducing the spread of epidemics using the 1918 Spanish Flu in the U.S. \cite{Hatchett2007-gc,Markel2007-ad,Bootsma2007-ww} and seasonal viral infections in France \cite{Adda2016-mn}. Our knowledge of economic impacts, however, is limited \cite{Wren-Lewis2020-vc}. For this question, past data may be less relevant, as the importance of face-to-face communication has increased steadily in the last 100 years through urbanization \cite{Henderson2010-mv,Henderson2002-ji} and specialization increased in business services as well \cite{Herrendorf2014-jx,Duarte2019-kc}. 
Even if advances in information and communication technology made it increasingly possible to communicate with co-workers and customers without the need for physical face-to-face interactions, personal contacts are still inevitable in some industries \cite{Dingel2020-lh,von2020labour}.

@@

To measure business disruptions from social distancing, we turn to recent data on the task descriptions of each occupation \cite{National_Center_for_ONET_Development2020-wj}, the precise geographic location of non-farm businesses in the U.S. \cite{CBP}, and customer mobility patterns \cite{SafeGraph2020-gn}. We construct three groups of occupations. First, some occupations require face-to-face communication several times a week with other workers. Examples of these \emph{teamwork-intensive} occupations include maintenance, personal care related occupations and health care professionals. Other occupations require frequent face-to-face contact with customers. Retail salespersons, social workers and waiters and waitresses are examples of such \emph{customer-facing} occupations. The third group of workers may need to be in physical proximity of one another even if they do not communicate, for example, to operate machinery or access key resources. Examples of such occupations requiring \emph{physical presence} are drivers and machine operators, especially in mining and water transport, where crammed working environments are common. With our occupation level measures, we aim to capture the jobs that can be performed less efficiently from home. We validate our indexes by using the American Time Use Survey (ATUS) \cite{ATUS}, which directly asks about the possibility of working from home.

To study how the patterns of interaction have changed in the U.S. during the Covid-19 pandemic, we use customer mobility data from SafeGraph \cite{SafeGraph2020-gn}. This dataset measures the number of visits to a business in a given month, as captured from several cell phone apps and made available to researchers in an anonymized form. We study how the reduced number of customer vists is correlated with changes in sectoral employment.

@@

The optimal number of contacts without social distancing is given by Eq \eqref{nstar}. Firms with $n^*>N$ are limited by social distancing. Their unit cost will increase to $c' = N \tau + N^{-\gamma}/\gamma$, which is greater than the optimal cost,
\begin{equation}\label{socdist}
\frac{c'}{c} 
	= \chi \frac{N}{n^*} + (1-\chi) \left(\frac{N}{n^*}\right)^{-\gamma}>1.
\end{equation}
The first term of the weighted average is less than one, representing a reduction in communication costs once the number of contacts declines. The second term is greater than one due to the fact that every worker has to complete a wider range of tasks than before, and they lose the benefit of specialization. Because $n^*$ is the cost-minimizing communication choice of the firm, the second term dominates, and production costs increase with social distancing.

@@

With our occupation-level measures, we aim to capture the type of activities that require frequent face-to-face contact. Our assumption is that these activities cannot be effectively done from home. To validate this assumption, we use data from the American Time Use Survey (ATUS) \cite{ATUS}, which directly asks workers whether they \emph{can} work from home. Fig \ref{customer_validation} plots our customer-intensity measure, for each occupation, against the share of workers who reported in ATUS that they can work from home (dark filled circles). Indeed, most customer-facing occupations have very few workers who can work from home. The pattern is different if we look at occupations that rely more on email, letters, and memos for customer communication (light hollow circles). The majority of these workers can work from home. Interestingly, for each degree of working from home (horizontal axis), there is sufficient variation in the importance of customer contact (vertical axis). The same patterns can be observed for teamwork-intensive occupations (Fig \ref{teamwork_validation}). This makes use conclude that dropping occupations primarily relying on email, letters and memos is sufficient to control for the potential to work from home.

@@

Using the cost change in Eq \ref{socdist}, we can express the compensating labor subsidy of industry $i$ in region $r$ as
\begin{equation}\label{subsidy}
    \lambda_{ir} = 1 -
        \frac{1-\chi_i}{1-\chi_i N_{ir}/n_{ir}^*}
        \left(
        \frac{N_{ir}}{n_{ir}^*}
        \right)^{\gamma_i}>0.
\end{equation}
We calibrate $n_{ir}^*$ to match the number of customer visits to establishments of industry $i$ in ZIP-code $r$ in February 2020, assuming that these visits were optimal before the pandemic hit the U.S. The new number of visits, $N_{it}$ will be calibrated to the number of customer visits in May 2020. The compensating wage subsidy increases in the importance of communication $\chi_i$ and decreases in the change in contacts $N_{ir}/n^*_{ir}$. The subscripts note that communication share is industry specific and the change in contacts is both industry and region specific.

@@

\begin{table}[!ht]
%\begin{adjustwidth}{-2.25in}{0in} % Comment out/remove adjustwidth environment if table fits in text column.
\caption{
{\bf Retail, accommodation and restaurants are the most communication intensive.}}
\begin{tabular}{|l+c|c|c+c|}
\hline
 & \multicolumn{3}{c}{\bf Communication} &  
\\ 
{\bf Industry} & {\bf Teamw.} & {\bf Custom.} & {\bf Overall} & {\bf Presence} 
\\ \thickhline
Retail trade 						&13	&66	&67	&5\\
Accommodation \& food services 		&8	&50	&51	&1\\
Arts, Entertainment, and Recreation &12	&38	&40	&2\\
Other Services (except Public Admin.)&12&30	&33	&12\\
Admin. \& Support \& Waste Manag.  	&17	&24	&27	&7\\
...&&&&\\
Wholesale Trade						&8	&12	&15	&12\\
Transportation and Warehousing 		&8	&8	&14	&33\\
Prof., Scient., and Technical Serv.	&5	&10	&12	&1\\
Manufacturing 						&7	&5	&9	&10\\
Agri., forestry, fishing \& hunting &4	&1	&4	&23\\

@@

We see three avenues for further research. The first concerns the interaction between sectors and regions. Whenever productivity in any business drops, this shock can propagate to its buyers and suppliers. The aggregate consequences of the epidemic will hence be modulated by input-output linkages between sectors, regions and countries \cite{Caliendo2014-mr,Caselli2020-nf,Baldwin2020-wb,Barrot2020-jf}.

The second and third directions concern the long-run response of businesses as they try to become more resilient to such shocks in the future. Whether the share of telecommunication remains large in the long run depends crucially on how easily it substitutes for face-to-face interaction. Previous work has found face-to-face communication to be more effective in high-intensity communication which is particularly helpful to overcome incentive problems in joint production \cite{Gaspar1998-gy,Storper2004-mg}. Data on internet flows suggests that telecommunication is not a good substitute for face-to-face meetings \cite{Cuberes2013-js}. None of these papers discuss disruptions from social distancing measures.

Third, businesses may change their location in response to perceived threats and disruptions. Epidemics have a disproportionate effect on cities. So it is conceivable that in a post-pandemic spatial equilibrium (not modeled here, but see \cite{Tian2019-wq}), the agglomeration premium falls and firms find it less attractive to locate in cities. A poignant point of comparison is the increased threat of terrorism in major cities following devastating attacks on New York, Washington, London, Paris, Madrid, Moscow and Mumbai. The general conclusion about terror threat is that cities have remained resilient and a robust attractor of businesses \cite{Glaeser2002-mw,Harrigan2002-ik}. We speculate that epidemics and social distancing can be more detrimental to cities than terror threats, because they tear apart the very fabric of urban life. However, we have limited data to make further predictions.

\section*{Data Availability Statement}
The mobility data used in this paper \cite{SafeGraph2020-gn} is confidential, but may be obtained free of charge for COVID-19-related research from the COVID-19 Consortium. Researchers interested in access to the data can apply at https://www.safegraph.com/covid-19-data-consortium. After signing a Data Agreement, access is granted within a few days. The authors will assist with any reasonable replication attempts for two years following publication.

@@

\section*{Acknowledgments}
We thank Gábor Békés, Péter Harasztosi, Péter Karádi, Balázs Lengyel, Dávid Krisztián Nagy, Andrea Weber, Charles Wyplosz, and two anonymous referees for comments.
